SMT-Boosted Security Types for Low-Level MPC

### 2025-01-30 esop25/camera-ready/intro.tex
checkpoint

Together, our contributions \emph{automate} the verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
\begin{enumerate}[i.]
\item In Section \ref{section-smt}, an interpretation of protocols in
  $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
  methodology for verifying correctness properties.
\item In Section \ref{section-cpj}, a confidentiality type system for
  $\minicat$ protocols with a soundness property guaranteeing that
  information is released only through explicit declassifications
  (Theorem \ref{theorem-cpj}).
\item In Section \ref{section-ipj}, an integrity type system for
  $\minicat$ protocols with a soundness property guaranteeing
  robustness in malicious settings (Theorem \ref{theorem-ipj}).
\item In Section \ref{section-metalang}, a dependent Hoare type system and
  algorithm for the $\metaprot$ language that is sound for the
  confidentiality and integrity type systems (Theorem
  \ref{theorem-mtj}), with independently verified, compositional $\Pi$
  types for functions.
\end{enumerate}
In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss extended example
applications of our type analyses to real protocols including
the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
(BDOZ), and Yao's Garbled Circuits (YGC) protocols.

@@

The goal of the $\minicat/\metaprot$ framework is to automate
correctness and security verification of low-level protocols.  Other
prior work has considered automated verification of high-level
protocols and manual verification of low-level protocols.
We summarize this comparison in Figure \ref{fig-comp-wrap}.

@@

\compwrapfig

>>>>

Together, our contributions \emph{automate} the verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
\begin{enumerate}[i.]
\item In Section \ref{section-smt}, an interpretation of protocols in
  $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
  methodology for verifying correctness properties.
\item In Section \ref{section-cpj}, a confidentiality type system for
  $\minicat$ protocols with a soundness property guaranteeing that
  information is released only through explicit declassifications
  (Theorem \ref{theorem-cpj}).
\item In Section \ref{section-ipj}, an integrity type system for
  $\minicat$ protocols with a soundness property guaranteeing
  robustness in malicious settings (Theorem \ref{theorem-ipj}).
\item In Section \ref{section-metalang}, a dependent Hoare type system and
  algorithm for the $\metaprot$ language that is sound for the
  confidentiality and integrity type systems (Theorem
  \ref{theorem-mtj}), with independently verified, compositional $\Pi$
  types for functions.
\end{enumerate}
In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss evaluation
and implementation of the the Goldreich-Micali-Wigderson (GMW)
protocol. In an extended version of this paper we also discuss
the Bendlin-Damgard-Orland-Zakarias
(BDOZ) protocol \cite{skalka2025smtboostedsecuritytypeslowlevel}.

@@

The goal of the $\minicat/\metaprot$ framework is to automate
correctness and security verification of low-level protocols.  Other
prior work has considered automated verification of high-level
protocols and manual verification of low-level protocols.
%We summarize this comparison in Figure \ref{fig-comp-wrap}.

@@

% \compwrapfig

### 2025-01-30 esop25/camera-ready/ipj.tex
checkpoint

In the above example, note that the environment implicitly records
that $\rvl{w}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. We explore
the application of this in more detail in Section
\ref{section-examples}.

@@

In all cases, we want to determine the integrity level of computed
values based on assumed partitionings $H,C$. In the previous example,
we want to say that $\rvl{w}$ has low integrity if either client 1 or
2 is corrupt.  To this end we define \emph{security labelings}
$\seclev$ which are mappings from variables $x$ to security levels
$\latel$. By default, given a partitioning $H,C$ this labeling maps
messages to the integrity level of the owner.
\begin{definition}  
  Given $H,C$,
  define $\seclev_{H,C}$ such that $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$
  if $\cid \in H$  and $\lolab$
  otherwise for all $\mx{w}{\cid}$.
\end{definition}
But the values of messages and reveals may be affected by values sent
by other parties, which is recorded in the $\Delta$ obtained from its
typing. To determine whether this impacts integrity for any given
$H,C$, we apply the inductively defined rewrite relation
$\cheatj{\Delta}{H,C}{\seclev}$, axiomatized in Figure
\ref{fig-cheatj}. The composition of typing and this latter rewriting
obtains essentially a traditional taint analysis, modulo MAC checking.
Our main type soundness result for integrity is then given as follows.
Proof details are given in the Appendix.
\begin{theorem}
  \label{theorem-ipj}
  Given  $\eqspre$ and $\prog$ with
  $\iov(\prog) = S \cup V \cup O$, if
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
  we have $\seclev(x) = \hilab$ for all $x \in \houtputs \cup O_H$, then $\prog$
  has integrity for all $H,C$.
\end{theorem}

>>>>

In the above example, note that the environment implicitly records
that $\rvl{w}$ depends on data from client 1, which in turn
depends on data from client 2. The rule for MAC
checking allows an overwrite of the dependency. An implementation
of a malicious-secure arithmetic circuit that
applies this rule is provided in extended version of
this paper \ref{skalka2025smtboostedsecuritytypeslowlevel}.

@@

In all cases, we want to determine the integrity level of computed
values based on assumed partitionings $H,C$. In the previous example,
we want to say that $\rvl{w}$ has low integrity if either client 1 or
2 is corrupt.  To this end we define \emph{security labelings}
$\seclev$ which are mappings from variables $x$ to security levels
$\latel$. By default, given a partitioning $H,C$ this labeling maps
messages to the integrity level of the owner.
\begin{definition}  
  Given $H,C$,
  define $\seclev_{H,C}$ such that $\seclev_{H,C}(\mx{w}{\cid} ) = \hilab$
  if $\cid \in H$  and $\lolab$
  otherwise for all $\mx{w}{\cid}$.
\end{definition}
But the values of messages and reveals may be affected by values sent
by other parties, which is recorded in the $\Delta$ obtained from its
typing. To determine whether this impacts integrity for any given
$H,C$, we apply the inductively defined rewrite relation
$\cheatj{\Delta}{H,C}{\seclev}$, axiomatized in Figure
\ref{fig-cheatj}. The composition of typing and this latter rewriting
obtains essentially a traditional taint analysis, modulo MAC checking.
Our main type soundness result for integrity is then given as follows.
Proof details are given in \cite{skalka2025smtboostedsecuritytypeslowlevel}.
\begin{theorem}
  \label{theorem-ipj}
  Given  $\eqspre$ and $\prog$ with
  $\iov(\prog) = S \cup V \cup O$, if
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and for all $H,C$ with $\cheatj{\Delta}{H,C}{\seclev}$ 
  we have $\seclev(x) = \hilab$ for all $x \in \houtputs \cup O_H$, then $\prog$
  has integrity for all $H,C$.
\end{theorem}

### 2025-01-30 esop25/camera-ready/metalang.tex
checkpoint

The $\metaprot$ language \cite{skalka-near-ppdp24} includes structured
data and function definitions for defining composable protocol
elements at a higher level of abstraction than $\minicat$.  The
$\metaprot$ language is a \emph{metalanguage} aka metaprogramming
language, where an $\fedprot$ protocol is the result of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
protocols. Separate verification of well-designed components results
in confidentiality and integrity properties to be recorded in their
types, allowing for significant reduction of SMT verification in whole
program analysis, as we will discuss with extended examples in Section
\ref{section-examples}.

@@

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which are guaranteed to
have closed types, to obtain Hoare typings for instructions. In Figure
\ref{fig-mtj} we define the type derivation rules for $\metaprot$. In
the \TirName{Mesg} rule we use algorithmic confidentiality and
integrity type checking in a straightforward manner.  In the
\TirName{Encode} rule, we verify the hint given in the type
annotation, and then use it for type checking.  In the \TirName{App}
rule we verify preconditions of the given function type and
instantiate the type and postconditions with the given expression
arguments.  The \TirName{Sig} rule we specify how to verify $\Pi$
types for functions. Note that the values $\mv_1,\ldots,\mv_n$ chosen
to instantiate the function parameters are chosen ``fresh'', meaning
that any strings and client identifiers occurring as substrings in
$\mv_1,\ldots,\mv_n$ are unique and not used in any program source
code. This ensures generality of verification as demonstrated in Lemma
\ref{lemma-eqs-notg} in the Appendix.

@@

We allow the programmer to specify a type signature $\tsig$ that
maps functions in $\codebase$ to valid dependent Hoare
typings. Define:
\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}
In practice, to define $\tsig$ the programmer only needs to specify
pre-~and postconditions, as we illustrate in Section
\ref{section-examples}, since the rest of the type is reconstructed by
type checking.

The following Theorem establishes our main result, that the
$\metaprot$ type checking is sound with respect to both
confidentiality and integrity typings. Since these typings imply
security hyperproperties of confidentiality and integrity, $\metaprot$
type checking implicitly enforces them. Proof details are given in the
Appendix.
\begin{theorem}[Soundness of $\metaprot$ type checking]
  \label{theorem-mtj}
  Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
  the judgement: $$\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$$ is derivable then
  $\instr \redx \prog$ and:
  \begin{enumerate}
  \item $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid.
  \item $\ipj{\eqs}{\prog}{\Delta}$ is valid.
  \end{enumerate}
\end{theorem}

>>>>

The $\metaprot$ language \cite{skalka-near-ppdp24} includes structured
data and function definitions for defining composable protocol
elements at a higher level of abstraction than $\minicat$.  The
$\metaprot$ language is a \emph{metalanguage} aka metaprogramming
language, where an $\fedprot$ protocol is the result of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
protocols. Separate verification of well-designed components results
in confidentiality and integrity properties to be recorded in their
types, allowing for significant reduction of SMT verification in whole
program analysis, as we illustrate with extended examples in Section
\ref{section-examples} and in an extended version of this paper
\cite{skalka2025smtboostedsecuritytypeslowlevel}.

@@

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which are guaranteed to
have closed types, to obtain Hoare typings for instructions. In Figure
\ref{fig-mtj} we define the type derivation rules for $\metaprot$. In
the \TirName{Mesg} rule we use algorithmic confidentiality and
integrity type checking in a straightforward manner.  In the
\TirName{Encode} rule, we verify the hint given in the type
annotation, and then use it for type checking.  In the \TirName{App}
rule we verify preconditions of the given function type and
instantiate the type and postconditions with the given expression
arguments.  The \TirName{Sig} rule we specify how to verify $\Pi$
types for functions. Note that the values $\mv_1,\ldots,\mv_n$ chosen
to instantiate the function parameters are chosen ``fresh'', meaning
that any strings and client identifiers occurring as substrings in
$\mv_1,\ldots,\mv_n$ are unique and not used in any program source
code. This ensures generality of verification as demonstrated in Lemma
\ref{lemma-eqs-notg} in an extended version of this paper \cite{skalka2025smtboostedsecuritytypeslowlevel}.

@@

We allow the programmer to specify a type signature $\tsig$ that
maps functions in $\codebase$ to valid dependent Hoare
typings. Define:
\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}
In practice, to define $\tsig$ the programmer only needs to specify
pre-~and postconditions, as we illustrate in Section
\ref{section-examples} and \cite{skalka2025smtboostedsecuritytypeslowlevel},
since the rest of the type is reconstructed by type checking.

The following Theorem establishes our main result, that the
$\metaprot$ type checking is sound with respect to both
confidentiality and integrity typings. Since these typings imply
security hyperproperties of confidentiality and integrity, $\metaprot$
type checking implicitly enforces them. Proof details are given \cite{skalka2025smtboostedsecuritytypeslowlevel}.
\begin{theorem}[Soundness of $\metaprot$ type checking]
  \label{theorem-mtj}
  Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
  the judgement: $$\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$$ is derivable then
  $\instr \redx \prog$ and:
  \begin{enumerate}
  \item $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid.
  \item $\ipj{\eqs}{\prog}{\Delta}$ is valid.
  \end{enumerate}
\end{theorem}
